# Exploring Parentage {#parents}

Now that I have my database all set up, I can explore the data.

```{r setup-db, eval = TRUE, echo = FALSE}
library(DBI)
library(tidyverse)

sheep <- dbConnect(drv = RSQLite::SQLite(), 
                   "../../../WILD6900_Comp-Tools/sheep.db")
individuals <- dbGetQuery(sheep, "SELECT * FROM individuals;")
fecal <- dbGetQuery(sheep, "SELECT * FROM fecal;")
census <- dbGetQuery(sheep, "SELECT * FROM census;")
group_locations <- dbGetQuery(sheep, "SELECT * FROM locations;")
group_composition <- dbGetQuery(sheep, "SELECT * FROM group_composition;")

```

I want to know who were the most successful ewes in reproductive fitness, so I'm going to use the mother_idno to track how many times it shows up for each ewe.

```{r mothers, eval = TRUE, echo = TRUE}
best_moms <- individuals %>% 
  select(name, mother_idno) %>%
  group_by(mother_idno) %>% 
  tally() %>%
  arrange(desc(n))

```
```{r top moms graph, eval = TRUE, echo = TRUE}

best_best_moms <- best_moms[c(2:21), ]

best_best_moms %>% 
  ggplot(mapping = aes(x = mother_idno, 
                       y = n,
                       fill = mother_idno)) +
  geom_col() + 
  labs(title = "Lambs by Mom",
    x = "mother id number",
    y = "number of lambs",
    fill = "mother_idno") +
  theme(axis.text.x = element_text(angle = 45, hjust=1))


```


I'm surprised that there were so many ewes over the years who had more than 10 lambs!

In the future, I would like to examine these data further and parse out how long those lambs lived. It could be that some of these ewes had many lambs, but only a few of those survived to adulthood. 
I would also like to take into account the family tree below each ewe to see how many grandchildren she had.

Because we have data for both parents of each animal, I can track the reproductive fitness of the rams too.

```{r fathers}
best_dads <- individuals %>% 
  filter(!is.na(father_idno)) %>% 
  select(name, father_idno) %>%
  group_by(father_idno) %>% 
  tally() %>%
  arrange(desc(n))

best_best_dads <- best_dads[c(2:21), ]

best_best_dads %>% 
  ggplot(mapping = aes(x = father_idno, 
                       y = n,
                       fill = father_idno)) +
  geom_col() + 
  labs(title = "Lambs by Dad",
    x = "father id number",
    fill = "father_idno") +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

```

One thing to note in this comparison is that there are 83 animals without data for a mother_idno, whereas there are 583 animals without data for a father_idno!

My instinct is that this is due to so many lambs over the years dying before they could be sampled for genetic material. It's much easier to record a mother for a lamb by observation alone.

The top father on this list is ram 00-03, who was the father of 21 lambs who survived long enough to sample for lineage. 

```{r id lookup, eval = TRUE, echo = TRUE}
individuals %>%
  select(id_number, name) %>% 
  filter(id_number == c("ID00-03", "ID00-02"))
```
Our best dad was Navel, and our best mom was Magpie. It's funny that they were both born in the same year, but I guess that was a good time period to be born into.
Now let's see who their children were.

```{r kids, eval = TRUE, echo = TRUE}
Magpies <- individuals %>%
  filter(mother_idno == "ID00-02")

Navels <- individuals %>%
  filter(father_idno == "ID00-03")
```

