# Group Size {#grp_size}

One of my main questions is about group size: whether there is a characteristic group size, when large groups form, and how typical group size scales with population. Large groups and flow between groups are important for disease transmission, and this dataset provides a rare opportunity to understand group sizes in bighorn sheep.

```{r group size, eval = TRUE, echo = TRUE}
#What was the biggest group observed and when?
census %>%
  as_tibble %>% 
  select(group_size, date) %>%
  group_by(group_size) %>% 
  arrange(desc(group_size))
  
```
It looks like the largest group was observed in June of 2014, and it was 85 individuals!

## Mode of Group size

I am curious about what group size is observed most frequently and what the average group size is for the whole dataset.

```{r mode group}
mode_fun <- function(x) {
     mode0 <- names(which.max(table(x)))
     if(is.numeric(x)) return(as.numeric(mode0))
     mode0
}
mode_fun(census$group_size)

```
The most common group size is 1! That is interesting to me. Now I want to see how often various group sizes occur, and if the smaller groups are more often observed than larger ones.

```{r common group sizes, echo=TRUE, eval=FALSE}
census %>% 
  as.tibble %>% 
  select(date, group_size) %>% 
  group_by(group_size) %>% 
  tally() %>% 
  arrange(desc(n))

```
Interesting!! The frequency of observed group sizes descends in order from 1 to 18! I expected there would be a group size that was more common, like 9 or 12. I also would have expected that our observational bias would be against the smaller groups because they would be easier to miss. I guess that's not so.

## Average Group Size

```{r overall avg, echo = TRUE, eval = FALSE}
mean(census$group_size)
```
This is the average group size over the entire dataset, which doesn't actually tell us that much. So. I'm going to look at how this has changed over the years.

```{r grps_by_year attempt, echo = TRUE, eval =FALSE}
library(lubridate)

grps_by_yr <- census %>% 
  as_tibble() %>% 
  mutate(ldate = mdy(census$date, tz = "US/Mountain")) %>% 
  select(ldate, group_size) %>% 
  group_by(year(ldate), group_size) %>% 
  tally() %>% 
  mutate(mean_per_year = mean(group_size)) %>% 
  filter(!is.na(ldate))
  rename(year = year(ldate))

grps_by_yr %>% 
  ggplot(mapping = aes(x = year(ldate), 
                       y = mean_per_year,
                       fill = year(ldate))) +
  geom_bar()
  
```


I want to know more about how group size is related to other factors like population size and sex ratio. To start exploring this, I want to look into when groups of different sizes occur and how that has changed over the 40 years in this dataset.
